---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading basic dependencies
```{r echo=TRUE}
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("dplyr")) install.packages("dplyr")
library("dplyr")
library("ggplot2")
```

## Loading and preprocessing the data
```{r echo=TRUE}
unzip(zipfile = "activity.zip")
activity <- read.csv(file = "activity.csv")
```

## What is mean total number of steps taken per day?
```{r}
daily.summary <- activity %>% group_by(date) %>% 
                            summarise(total.steps = sum(steps,na.rm = T))
ggplot(data=daily.summary, aes(total.steps)) + geom_histogram() 
```

The average number of steps taken daily is `r round(mean(daily.summary$total.steps))`.

## What is the average daily activity pattern?
```{r echo=TRUE}
interval.summary <- activity %>% group_by(interval) %>% 
                              summarise(avg.steps = mean(steps, na.rm = T)) %>% 
                              arrange(desc(avg.steps))

ggplot(data= interval.summary, aes(interval, avg.steps)) + geom_line()
```

The interval in the day `r head(interval.summary$interval,1)`.

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
